---
title: "CS_03_answers"
author: "Reed"
date: "1/24/2022"
output: pdf_document
---


```{r}
library(tidyverse)
library(lubridate)
library(here)
p <- 'C:/R_datsci_2022/data'
f <- list.files(p, 'csv')
fp <- paste0(p, '/', f)

census_size <- read.csv(fp[2])
traits <- read.csv(fp[3])

rm(f, p, fp)
```

```{r}
rare_plants <- left_join(census_size, traits, by = c('species' = 'species_name'))

rm(census_size, traits)
```

Q1. For each species, how many populations are featured in this dataset?

```{r}

rare_plants %>% 
  group_by(species) %>% 
  distinct(species, pop_id) %>% 
  tally() # use tally

rare_plants %>% 
  group_by(species) %>% 
  distinct(species, pop_id) %>%  
  count() # or use count. 

rare_plants %>% 
  group_by(species) %>% 
  distinct(species, pop_id) %>% 
  nrow() # you cannot use nrow, it is from base r and does
## not recognize groupings

```

Q2. Which population has the most years of data available (biggest sample size)?

```{r}
rare_plants %>% 
  ungroup() %>% 
  #drop_na(pop_size) %>% 
  #group_by(species, pop_id) %>% 
  #tally() %>% 
  count(species, pop_id) %>% 
  arrange(-n) %>% 
  slice(2, with_ties = T) # weird needs slice for both
```

Q3. Do annual or perennial species have higher population counts on average? What about variance? Hint: use the appropriate join function to join the tables. Then us group_by(), and finally the mean() and var() functions in summarize().

```{r}

rare_plants %>% 
  group_by(duration) %>% 
  summarize(Average_Count = mean(pop_size, na.rm = T),
            Average_Variance = var(pop_size, na.rm = T))

```

```{r}
rare_plants$date <- mdy(rare_plants$date)

lvls <- rare_plants %>% group_by(species) %>% 
  distinct(pop_id) %>% 
  arrange(species) %>% 
  mutate(pop_id = as_factor(pop_id)) %>% 
  pull(pop_id)

rare_plants <- rare_plants %>% mutate(Population_id = as_factor())

rare_plants %>% 
  filter(complete.cases(.)) %>% 
  mutate(pop_id_char = as.character(pop_id)) %>% 
  #mutate(pop_id_fact = as_factor(pop_id_char)) %>% 
  #mutate(pop_id_fact = fct_reorder(pop_id_fact, lvls)) %>% 
  mutate(species = str_to_sentence(species)) %>% 
  ggplot(aes(date, log(pop_size), color = pop_id_char)) + 
  theme_bw() +
  guides(color = guide_legend("Population ID"))  +
  labs(x = 'Year', y = 'Number of Individuals (log)', title = 'Trends in Poplation Sizes Overtime') +
  theme(plot.title = element_text(hjust = 0.5))+
  geom_line() + 
  geom_point() + 
  facet_wrap(~ species) 

rare_plants %>% 
  filter(complete.cases(.)) %>% 
  mutate(pop_id_char = as.character(pop_id)) %>% 
  mutate(species = str_to_sentence(species)) %>% 
  
  ggplot(aes(date, pop_size, color = pop_id_char)) + 
  theme_bw() +
  guides(color = guide_legend("Population ID"))  +
  labs(x = 'Year', y = 'Number of Individuals', title = 'Trends in Poplation Sizes Overtime') +
  theme(plot.title = element_text(hjust = 0.5))+
  geom_line() + 
  geom_point() + 
  facet_wrap(~ species) 
```
