---
title: "CS_01_Iris_Flowers"
author: "Reed"
date: "1/12/2022"
output: pdf_document
---
Case Study One.

Iris flowers statistics

Load these data and take a look at them. 
```{r}
data(iris)
head(iris)
```

What species are in the data set?
```{r}
unique(iris$Species)
```

How many species are in the data set?
```{r}
length(unique(iris$Species))
```

# Main section 


A one off calculation
```{r}
var<- iris[iris$Species=="setosa",]
mean(var$Sepal.Length)
```


# Aggregate

In this example we calculate the *Median* of *Sepal Width* for each species. 

Here we specify the data which contains our values of interest. We specify that we want to calculate the median of the value Sepal Width for by (~) each Species in the data.
```{r}
Median <- aggregate(iris$Sepal.Width ~ iris$Species, data = iris, 
          FUN = function(x) {median(x)})

colnames(Median) <- c('Species', 'Value')
```


In this example we find the *Maximum* value of *Petal Length* for each species. 

We can use aggegate, after we use list to split the dataframe into several dataframes each containing only one species.

```{r}
Max <- aggregate(iris$Petal.Length, list(iris$Species), FUN=max) 

colnames(Max) <- c('Species', 'Value')
```


# Using the apply family

## **t**apply

In this example we calculate the *Mean* of *Sepal length* for each species.
```{r}
Mean <- as.data.frame(
  tapply(iris$Sepal.Length, iris$Species, mean)
)

Mean <- cbind(rownames(Mean), Mean)
colnames(Mean) <- c('Species', 'Value')
rownames(Mean) <- NULL
```

## **L**apply

in this example we find the *Minimum* value of *Petal Length* for each species.
```{r}
iris_list <- split(iris, iris$Species)
Min <- lapply(iris_list, function(x) min(x$Petal.Length))
Min <- t(as.data.frame(Min))
Min <- as.data.frame(cbind(rownames(Min), Min))
colnames(Min) <- c('Species', 'Value')
rownames(Min) <- NULL

rm(iris_list)
```

In this example we calculate the *Mean* of each trait for a single species.
```{r}
iris_setosa <- iris[iris$Species == 'setosa', ]
iris_setosa <- subset(iris,select =  -Species)

setosa_mean <- as.data.frame(unlist(lapply(iris_setosa, mean)))
```

in this example we calculate each *Summary* statistic for *all* species.... this is what happens when no one shows up for an office hour...
```{r}

iris_summary <- function(x){
  
  Species <- as.character(x$Species)
  x[,1:4]
  data_summary <- as.data.frame(unlist(lapply(x, summary)))
  data_summary <- as.data.frame(cbind(rownames(data_summary), data_summary))
  data_summary <- setNames(data_summary, c('Statistic', 'Value'))
  data_summary <- data_summary[1:24,]
  
  data_summary$Statistic <- as.character(data_summary$Statistic)
  variables <- strsplit(data_summary$Statistic, "//.")
  variables <- t(as.data.frame(variables))
  variables <- as.data.frame(variables)
  variables <- setNames(variables, c('Tissue','Dimension','Statistic'))
  Value <- data_summary[,2]

  Species <- Species[1:nrow(variables)]
  results <- cbind(Species, variables, Value)
  rownames(results) <- NULL
  return(results)

}


iris_list <- split(iris, f = iris$Species)
spp_stats <- lapply(iris_list, iris_summary)
spp_stats <- do.call(rbind, spp_stats)
rownames(spp_stats) <- NULL
spp_stats

rm(iris_list, iris_summary)

```


# Create One Data Frame of the Summary Stats

We will create a 'Long' Dataframe, where each value is described attributes in the adjacent columns. 

```{r}

Mean$Statistic <- 'Mean'
Median$Statistic <- 'Median'
Min$Statistic <- 'Min'
Max$Statistic <- 'Max'

Mean$Feature <- 'Sepal.Length'
Median$Feature <- 'Sepal.Width'
Min$Feature <- 'Petal.Length'
Max$Feature <- 'Petal.Length'

dataframe <- rbind(Mean, Median, Min, Max)

rm(Mean, Median, Min, Max)
```

  
# Plot Histograms of Petal Length for every species. 

```{r}
par(mfrow = c(3,1))

hist(iris$Petal.Length[iris$Species=='setosa'], 
     main = 'Iris setosa Pall. ex Link',
     xlim = c(1, 7), 
     xlab = "",
     density = 50, 
     angle = 45,
     col = "purple", 
     border = "black")
hist(iris$Petal.Length[iris$Species=='versicolor'], 
     main = 'I. versicolor L.', 
     xlim = c(1,7), 
     xlab = "", 
     density = 50, 
     angle = 45,
     col = "purple", 
     border = "black")
hist(iris$Petal.Length[iris$Species=='virginica'], 
     main = 'I. virginica L.', 
     xlim = c(1,7), 
     xlab = "Length in cm", 
     density = 50, 
     angle = 45,
     col = "purple", 
     border = "black")

```


