---
title: "CS_02_Answers"
author: "steppe"
date: "1/16/2022"
output: html_document
---


```{r load libraries, message  = F, warning = F}
library(gapminder)
library(tidyverse)
library(data.table) # tidyverse alternative "strictly for linemen not for freshmen"
```


```{r}
gapminder
gapminder <- as.data.frame(gapminder)
head(gapminder)

```


Remove Kuwait
```{r Remove Kuwait}

# TIDYVERSE
gapminder_tv <- filter(gapminder, country != 'Kuwait') # option 1
gapminder_tv <- gapminder %>% filter(!country %in% 'Kuwait') # option 2

'%notin%' <- Negate('%in%')
to_rm <- c('Kuwait') # better for lists.
gapminder_tv <- gapminder %>% filter(country %notin% to_rm) # option 3 - weird

# BASE R
gapminder_b <- gapminder[gapminder$country != 'Kuwait',] # option 1
gapminder_b <- gapminder[which(gapminder$country!='Kuwait'),] # option 2
gapminder_b <- base::subset(gapminder, country != 'Kuwait') # option 3

# DATA.TABLE
gapminder_dt <- setDT(gapminder) # important. 
gapminder_dt <- gapminder_dt[country != 'Kuwait']

rm('%notin%', to_rm)
```


• Q1. What are the top three countries with the *highest* or the *lowest* life expectancy at birth in 1957
and 2007?

```{r Life Expectancy top bottom three}

# TIDYVERSE
life_tv <- gapminder_tv %>% # option 1,2,3 & 4
  filter(year == 1957 | year == 2007) %>% # likewise year %in% c(1957, 2007) works
  group_by(year) %>% 
  arrange(-lifeExp) %>%  # -
  # ARRANGE defaults to ascending order,
  # we can remove the '-' and return the longest lifes this way to
  slice_head(n = 3)
  # for each pair of '-' the the slice can be switched.

# BASE R
# 4 combinations here. 
life_b_57 <- gapminder_b[gapminder_b$year == 1957,] #  - this can be changed for '2007'
life_b_57 <- life_b_57[order(life_b_57$lifeExp, decreasing = TRUE), ] # this can be flipped T > F
life_b_57 <- base::Reduce(rbind, by(life_b_57, life_b_57["year"], head,n = 3)) # 'head' can be flipped to 'tail'
life_b_57 <- life_b_57[1:3,] # by hand for both ways... also life_b_57[nrows(life_b_57):nrows(life_b_57)-3]
# life_b_57 <- life_b_57[(nrow(life_b_57)-2):nrow(life_b_57),] # a gross complement to the above option [1:3]


# DATA.TABLE
# 4 combinations here. 
life_dt <- gapminder_dt[year == 1957 | year == 2007][order(lifeExp), head(.SD, 3), by = "year"]
# likewise year %in% c(1957, 2007) works 
# can flip to 'tail.' 

rm(life_tv, life_b_57, life_dt)
```

• Q2. How many countries had life expectancy at birth smaller than 50 in 1957 and 2007?

```{r}

# TIDYVERSE
life_exp_tv <- gapminder_tv %>% # works
  filter(year == 1957 & lifeExp < 50 | year == 2007 & lifeExp < 50) %>% 
  group_by(country) %>% 
  count() %>% 
  filter(n >= 2) %>% # should be 18...
  nrow()

# still in works BASE R
life_exp_b <- with(gapminder_b, gapminder_b[(year==1957 & lifeExp < 50) | (year==2007 & lifeExp < 50), ])
counts <- table(life_exp_b$country)
life_exp_b <- life_exp_b[life_exp_b$country %in% names(counts[counts == 2]), ]
length(unique(as.character(life_exp_b$country)))

# DATA.TABLE
# option 1
life_exp_dt <- gapminder_dt[year %in% c(1957,2007) & lifeExp < 50][, counts := .N, by = country][counts >= 2]
count(unique(life_exp_dt, by = 'country'))

# option 2
life_exp_dt <- gapminder_dt[year %in% c(1957,2007) & lifeExp < 50][, if (.N == 2) .SD, by = country]
count(unique(life_exp_dt, by = 'country'))

rm(life_exp_tv, life_exp_b, life_exp_dt)
```

• Q3. What are total population in 1957 and 2007, in Asia and Americas respectively?

```{r}
# TIDYVERSE
total_pop_tv <- gapminder_tv %>% 
  filter(continent %in% c('Asia', 'Americas') & year %in% c(1957, 2007)) %>% 
  group_by(continent, year) %>% 
  summarise(.groups = "keep", sum(pop))

#with(gapminder_b, gapminder_b[(which(year %in% c(1957, 2007))) & which(continent %in% c('Asia', 'Americas'))])

# DATA.TABLE
total_pop_dt <- gapminder_dt[continent %in% c('Asia', 'Americas') & year %in% c(1957, 2007)][,new := sum(pop), by = list(continent, year)]
  
rm(total_pop_tv)
```

• Q4. How many countries had population higher than 30,000,000 in 1957 and 2007?

```{r}
# TIDYVERSE
nrow(gapminder_tv %>% filter(year == 1957 & pop > 30000000))
nrow(gapminder_tv %>% filter(year == 2007 & pop > 30000000))

# BASE 
nrow(with(gapminder_b, gapminder_b[(year==1957 & pop > 30000000), ]))
nrow(with(gapminder_b, gapminder_b[(year==2007 & pop > 30000000), ]))

# DATA.TABLE
nrow(gapminder_dt[year == 1957 & pop > 30000000])
nrow(gapminder_dt[year == 2007 & pop > 30000000])
```

• Q5. What are the top five countries with the highest total GDP (pop*gdpPercap) in 1957 and 2007?

```{r}

# TIDYVERSE
t5_Tgdp_tv <- gapminder_tv %>% 
  filter(year %in% c(1957, 2007)) %>% 
  mutate('Tgdp' = pop*gdpPercap) %>% 
  group_by(year) %>% 
  slice_max(Tgdp, n = 5) 

# BASE
t5_Tgdp_b <- gapminder[gapminder$year %in% c(1957, 2007),]
t5_Tgdp_b$Tgdp <- t5_Tgdp_b$pop * t5_Tgdp_b$gdpPercap
t5_Tgdp_b <- t5_Tgdp_b[order(t5_Tgdp_b$year, t5_Tgdp_b$Tgdp, decreasing = T),]
t5_Tgdp_b <- split(t5_Tgdp_b, f = t5_Tgdp_b$year)
t5_Tgdp_b <- lapply(t5_Tgdp_b, function(x) head(x, n = 5))
t5_Tgdp_b <- do.call("rbind", t5_Tgdp_b)

# DATA.TABLE
t5_Tgdp_dt <- gapminder_dt[year %in% c(1957,2007)][,Tgdp := pop*gdpPercap][order(Tgdp), tail(.SD, 5), by = year]

rm(t5_Tgdp_tv, t5_Tgdp_b, t5_Tgdp_dt)
```

• Q6. What are the mean values of lifeExp and gdpPercap in 1957 and 2007, for Africa and Europe
respectively?

```{r}
# TIDYVERSE
means_tv <- gapminder_tv %>% 
  filter(year %in% c(1957, 2007) & continent %in% c('Africa', 'Europe')) %>% 
  group_by(continent, year) %>% 
  mutate(mean_LE = mean(lifeExp)) %>% 
  mutate(mean_GDP = mean(gdpPercap)) %>% 
  dplyr::select(continent, year, mean_LE, mean_GDP) %>% 
  distinct()

# BASE
means_b <- with(gapminder_b, gapminder_b[(year %in% c(1957, 2007) & continent %in% c('Africa', 'Europe')),])
means_b <- aggregate(. ~ means_b$continent + means_b$year, data = means_b, 
          FUN = function(x) {mean(x)})
means_b <- means_b[,c(1,5:6,8)]
colnames(means_b) <- c('continent','year','mean_LE','mean_GDP')

# DATA.TABLE
grp_vars <- c('continent','year')
means_dt <- gapminder_dt[continent %in% c('Africa', 'Europe') & year %in% c(1957, 2007)][,mean_LE := mean(lifeExp), by = grp_vars][,mean_GDP := mean(gdpPercap), by = grp_vars][, c("continent","year","mean_LE","mean_GDP")]
setkeyv(means_dt, grp_vars)
means_dt <- subset(unique(means_dt))

rm(means_tv, means_b, grp_vars, means_dt)
```


Scatterplots of lifeExp, pop and gdpPercap of United States over the years from 1952
to 2007.


```{r Scatterplots}

# TIDYVERSE

sp_data <- filter(gapminder_tv, country == 'United States') %>% mutate(pop = pop/1000000) 

a <- ggplot(sp_data, aes(x = year, y = lifeExp)) +
  geom_point() +
  theme_classic() +
  labs(x = "Year", y = "Life Expectancy",
       title = "Life Expectancy in the United States (1952-2007)") +
  theme(plot.title = element_text(hjust = 0.5)) 
b <- ggplot(sp_data, aes(x = year, y = pop)) +
  geom_point() +
  theme_classic() +
  labs(x = "Year", y = "Population in Millions",
       title = "Population in the United States (1952-2007)") +
  theme(plot.title = element_text(hjust = 0.5)) 
c <- ggplot(sp_data, aes(x = year, y = gdpPercap)) +
  geom_point() +
  theme_classic() +
  labs(x = "Year", y = "GDP per capita",
       title = "Gross Domestic Product (GDP) per Capita in the United States (1952-2007)") +
  theme(plot.title = element_text(hjust = 0.5)) 


d <- cowplot::plot_grid(a,b,c, labels = NULL, ncol = 1)
ggsave("US_Scatterplot_tv.png", plot = d, dpi = 300)

# Base R
sp_data <- gapminder_b[gapminder_b$country == "United States",]
sp_data$pop <- sp_data$pop/1000000

png(file = "US_Scatterplot_base.png", width = 900, height = 600)
par(mfrow = c(3,1))
plot(x=sp_data$year, y=sp_data$lifeExp, xlab = "Year", ylab = "Life Expectancy", pch = 19,
     main = "Life Expectancy in the United States (1952-2007)")
plot(x=sp_data$year, y=sp_data$pop, xlab = "Year", ylab = "Population in Millions", pch = 19,
     main = "Population in the United States (1952-2007)")
plot(x=sp_data$year, y=sp_data$gdpPercap, xlab = "Year", ylab = "GDP per capita", pch = 19,
     main = "Gross Domestic Product (GDP) per Capita in the United States (1952-2007)")

dev.off()

rm(sp_data, a, b, c, d)
```

