---
title: "Emails"
author: "Reed"
date: "3/14/2021"
output: pdf_document
---
 
```{r, loadlib, results='hide', message=F, warning=F}
library(tidyverse)
```

## Emails
Many students emailed questions in addition to attending office hours.

This is not the greatest CSV format ever (it was rendered by outlook), and scrubbing the info would take more time than I think is required for this purpose. But, here is a way of tracking how many emails I wrote. 
```{r, Emails wrangling}
log <- read.csv("TA_EMAILS_LOG.csv")[,c(1:4,6:7)] %>% #neccessary columns
  janitor::clean_names()

log <- log %>% 
  mutate(body = str_replace_all(body, "[\r\n\t]" , " ")) %>% # gross
  filter(from_address == 'reedbenkendorf2021@u.northwestern.edu') %>% #only emails I wrote
  filter(to_address != "yingying.xie@northwestern.edu") %>% # i didn't write you any long emails
  mutate(meeting_link = str_detect(body, "Reed Benkendorf is inviting you to a scheduled Zoom meeting.")) %>% # these were all short emails were people needed links
  filter(meeting_link == FALSE) %>% #remove them now
  dplyr::select(-meeting_link) %>% #don't need the column anymore
  mutate(body = str_remove_all(body, '<.*>')) %>% #remove tags and contents within them to...
  mutate(num_char = nchar(body)) %>% #arrange the length of the messages to..
  filter(num_char > 262) # delete little short follow up messages

emails_sent <- nrow(log)
```

I estimate the mean time spent troubleshooting a problem and writing an email was 10 minutes. 
```{r, Emails final, comment=""}
total_email_time <- c('Time spent responding to emails:' =  (emails_sent * 10)/60) # minutes to hour
writeLines(paste(names(total_email_time), total_email_time, "hours", sep = " ", collapse = ","))

hours_remain <- hours_remain - total_email_time

rm(log, emails_sent, total_email_time)
```
 
 
 ## Time spend writing the script for attendance and downloading the CSV's
 
 The attendance script took some time, this one only took about an hour. 
```{r, comment=""}
scripts <- c('Total time spent working on scripts:' = 5)
writeLines(paste(names(scripts), scripts, "hours", sep = " ", collapse = ","))

hours_remain <- hours_remain - scripts
rm(scripts)
```

## Total time worked

```{r, comment=""}
hours_worked <- round(total_hours - hours_remain, 2)

writeLines(paste("I worked", hours_worked, "hours"))
```

